100 Years of Social Work Research: A Data Science Perspective

### 2015-01-09 Analysis.Rmd
Updated search results. Minor code modifcation.  Funtioning.

# Overview of data

The original data were from a search of PsychInfo using Ebsco Host platform (December 23, 2014). The following search operators and limiters were used:

+ SO "social work" OR SO "social welfare" OR SO "social casework" OR SO "social services"

+ Limiters - Document Type: Journal Article

+ Search modes - Boolean/Phrase	Interface - EBSCOhost Research Databases

+ Search Screen - Advanced Search

+ Database - PsycINFO

The search results were exported in a _generic bibliographic format_, which is an unstructured text (*.txt) file.The text file was processed using the `BibWrangleR` function created by the first author.

# Initialize OS-X workspace and functions for data wrangling

This section processes raw data.  This section of code is executed only one time to transform raw text data into an analyzable format.  When new data are obtained for this study (i.e., updated search results), this section should be re-run by changing `echo=FALSE` to `echo=TRUE` in the knitr markdown argument.

@@

# Read BWR functions for Mac OS
source("/Users/beperron/Git/BibWrangleR/functions/piWrangleR.R")
source("/Users/beperron/Git/BibWrangleR/functions/packages.R")
# Set the path where original raw data are stored
setwd("/Users/beperron/Git/SocialWorkResearch")

# Set the working directory to store files created by BWR functions
my.path <- "/Users/beperron/Git/SocialWorkResearch"

# Wrangle the data with the BWR function suite
#piBWR.f(csv=FALSE, path=my.path)
#save(pi.df, file = "piArticles.R")
```

# Initialize Windows workspace and functions for data wrangling

# Initialize workspace and functions for analaysis

All the analyses performed involve the data that have been processed with the `BibWrangleR` functions.  This section reads the processed data, loads the required packages, and does a quick quality check to ensure that the same number of articles (i.e., records) contained in the original search match the number of articles in the transformed data.

```{r initialization_analysis, message=FALSE, warning=FALSE, comment=NA}
rm(list=ls())
setwd("/Users/beperron/Git/SocialWorkResearch")
source("/Users/beperron/Git/BibWrangleR/functions/ggsurv.R")
load("piArticles.R")
library(dplyr)
library(ggplot2)
library(gridExtra)
library(survival)
library(grid)
library(png)

@@

# How many unique article titles?  Ebsco Results of most current search is $n=24,314$. Do not proceed with analysis if the output does not match this result.  
length(which(pi.df$attributes == "TI"))
```

## What is the overall number and names of journal titles?

```{r unique_titles}
unique.titles <- filter(pi.df, attributes == "SO")

# Number of unique titles
length(unique(unique.titles$record))

# Unique titles
unique(unique.titles$record)
```

@@

```{r unique_titles_year, message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4}
journals.year <- tbl_df(pi.df)

@@

```{r journal_output}

@@

```{r journal_lifespan}
#10 longest running journals 
longest.running <- n.so.yr %>%
       mutate(last = as.numeric(last), first = as.numeric(first), 
              year.diff = last - first) %>%
       arrange(desc(year.diff)) %>%
       select(title, first, last, year.diff) %>%
       mutate(stop = year.diff, event = ifelse(as.numeric(last) != 2014, 1, 0)) %>%
       select(title, stop, event, as.numeric(first))

@@

ggsurv(survival.journals) + 
    geom_line(data = median.survival, aes(time, quant), linetype="longdash") +
    annotate("segment", x = 18, xend = 12, y = .12, yend = .15, size = .25, arrow =arrow()) +
    geom_text(x = 29, y =.12, label = "median survival", size = 4) + 
    geom_text(x = 75, y =.75, label = "+ = Last year in print", size = 4)

@@

```{r articles_per_year}
n.articles.year <- filter(pi.df, attributes == "YR") 
year.split <- split(n.articles.year, n.articles.year$record)
year.count <- unlist(lapply(year.split, nrow))
year.count <- year.count[order(names(year.count))]
years <- names(year.count)

@@

# Print most recent ten years
head(df, 10)
```

# What are the topic areas (by Subject Terms)?

```{r subject_terms_overall, comment=NA, warning=FALSE, message=FALSE}

su.df <- filter(pi.df, attributes == "SU")

subject.terms <- stringr::str_split(su.df$record, pattern = ";")
subject.terms <- unlist(lapply(subject.terms, function(x) gsub(" ", "", x)))

subject.terms.total <- length(unlist(lapply(subject.terms, 
                        function(x) gsub(" ", "", x))))

# Print 25 most commmon terms
head(most.frequent, 25)

## What are the topic areas over time (by Subject terms)?

```{r subject_terms_overtime, warning=FALSE, message=FALSE}
decade <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

decade$year <- cut(decade$year, breaks = 10, labels = c(1:10))

keywords <- pi.df %>%
        filter(attributes == "SU") %>%
        select(articleID = articleID, keywords = record)

keywords.decade <- keywords %>% 
         left_join(decade)

library(plyr)
keywords.data.split <- dlply(keywords.decade, .(year))
detach(package:plyr)

terms.f <- function(x){
    split.terms <- stringr::str_split(x[,"keywords"], pattern =";")
    clean.terms <- lapply(split.terms, function(x) gsub(" ", "", x))
    }

keywords.decade <- lapply(keywords.data.split, terms.f)
keywords.decade <- lapply(keywords.decade, unlist)

temp <- lapply(keywords.decade, function(x) data.frame(table(x)))
temp <- lapply(temp, function(x) arrange(x, desc(Freq)))
lapply(temp, function(x) head(x,10))    
```

@@

```{r keywords_overall, comment=NA, warning=FALSE, message=FALSE}
kp.df <- filter(pi.df, attributes == "KP")

@@

```{r keywords_over_time, warning=FALSE, message=FALSE}
decade <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

decade$year <- cut(decade$year, breaks = 10, labels = c(1:10))

keywords <- pi.df %>%
        filter(attributes == "KP") %>%
        select(articleID = articleID, keywords = record)

keywords.decade <- keywords %>% 
         left_join(decade)

library(plyr)
keywords.data.split <- dlply(keywords.decade, .(year))
detach(package:plyr)

terms.f <- function(x){
    split.terms <- stringr::str_split(x[,"keywords"], pattern =";")
    clean.terms <- lapply(split.terms, function(x) gsub(" ", "", x))
    }

keywords.decade <- lapply(keywords.data.split, terms.f)
keywords.decade <- lapply(keywords.decade, unlist)

temp <- lapply(keywords.decade, function(x) data.frame(table(x)))
temp <- lapply(temp, function(x) arrange(x, desc(Freq)))
lapply(temp, function(x) head(x,10))    
```

@@

```{r location, comment=NA, warning=FALSE, message=FALSE}

@@

print(location)

img2 <- readPNG("/Users/beperron/Git/SocialWorkResearch/Chloro.png")
grid.raster(img2)
```

@@

```{r location_over_time, warning=FALSE, message=FALSE}

@@

This section is currently not functional
```{r countries_not_represented}
#Tableau.countries <- read.csv("/Users/beperron/Git/SocialWorkResearch/SupportingDocs/Tableau\ Countries.csv")
```

@@

```{r methodology, comment=NA, warning=FALSE, message=FALSE}

@@

```{r}
year <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

>>>>

# Overview of Search Strategy

Date conducted:   January 8, 2014

Database: PsychInfo (via EbscoHost Platform)

Search terms:  SO "social work" or SO "social welfare" or SO "social casework" or "social case work" or SO "social service" or "social services" or SO "human service" or SO "human services" or SO "social development" or SO "social environment"

Search limiters:  Publication Type: Peer Reviewed Journal; Document Type: Journal Article

Search modes: Boolean/Phrase

Search results were exported from EbscoHost Platformon in a _generic bibliographic format_.  This is an unstructured text file (*.txt) that was processed using in R.

@@

# Read BWR functions for Mac OS
source("/Users/beperron/Git/BibWrangleR/functions/piWrangleR.R")
source("/Users/beperron/Git/BibWrangleR/functions/packages.R")
# Set the path where original raw data are stored
setwd("/Users/beperron/Git/SocialWorkResearch/Data")

# Set the working directory to store files created by BWR functions
path <- "/Users/beperron/Git/SocialWorkResearch/Data"

# Wrangle the data with the BWR function suite
#piBWR.f(csv=FALSE, path=path)
#save(pi.df, file = "piArticles.R")
```

```{r initialization_analysis, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
rm(list=ls())
setwd("/Users/beperron/Git/SocialWorkResearch/Data")
source("/Users/beperron/Git/BibWrangleR/functions/ggsurv.R")
load("piArticles.R")
library(dplyr)
library(ggplot2)
library(gridExtra)
library(survival)
library(grid)
library(png)

@@

# How many unique article titles?  Ebsco Results of most current search is $n=XXXX$. Do not proceed with analysis if the output does not match this result.  
length(which(pi.df$attributes == "TI"))
```

```{r echo=FALSE}
pi.df <- pi.df %>% filter(
        record != "Journal of Applied Social Sciences" &
        record != "Early Child Development and Care" &
        record != "The Clinical Supervisor" &
        record != "Children and Youth Services Review" &
        record != "General Hospital Psychiatry" &
        record != "Canadian Journal on Aging" &
        record != "Canadian Journal of Community Mental Health" &
        record != "Behavior Modification" &
        record != "Employee Assistance Quarterly" &
        record != "Journal of Applied Behavioral Science" & 
        record != "The Scientific Review of Mental Health Practice: Objective Investigations of Controversial and Unorthodox Claims in Clinical Psychology, Psychiatry, and Social Work" &
        record != "Systems Research and Behavioral Science" & 
        record != "Journal of Applied Rehabilitation Counseling" & 
        record != "Journal of Educational & Psychological Consultation" &
        record != "Journal of Community & Applied Social Psychology" &                                          
        record != "Evaluation and Program Planning" &                                                                  
        record != "Journal of Nonverbal Behavior" &
        record != "Brain and Cognition" &
        record != "Early Education and Development")
```

# What is the overall number and names of journal titles?

```{r unique_titles, echo=FALSE}
unique.titles <- filter(pi.df, attributes == "SO")

# Number of unique titles
number.journals <- length(unique(unique.titles$record))

# Unique titles
journals.unique <- unique(unique.titles$record)
#write.csv(journals.unique, "journals.csv")

number.journals
journals.unique
```

@@

```{r unique_titles_year, message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4, echo=FALSE}
journals.year <- tbl_df(pi.df)

@@

```{r journal_output, echo=FALSE}

@@

```{r journal_lifespan, echo=FALSE}
#10 longest running journals 
longest.running <- n.so.yr %>%
       mutate(last = as.numeric(last), first = as.numeric(first), 
              year.diff = last - first) %>%
       arrange(desc(year.diff)) %>%
       select(title, first, last, year.diff) %>%
       mutate(stop = year.diff, event = ifelse(as.numeric(last) != 2014, 1, 0)) %>%
       select(title, stop, event, as.numeric(first))

@@

ggsurv(survival.journals) + 
    #geom_line(data = median.survival, aes(time, quant), linetype="longdash") +
    #annotate("segment", x = 18, xend = 12, y = .12, yend = .15, size = .25, arrow =arrow()) +
    #geom_text(x = 29, y =.12, label = "median survival", size = 4) + 
    geom_text(x = 75, y =.75, label = "+ = Last year in print", size = 4) +
    ylim(0,1)

@@

```{r articles_per_year, echo=FALSE}
n.articles.year <- filter(pi.df, attributes == "YR") 
year.split <- split(n.articles.year, n.articles.year$record)
year.count <- unlist(lapply(year.split, nrow))
year.count <- year.count[order(names(year.count))]
years <- names(year.count)

@@

# Print most recent ten years
tail(df, 10)
```

@@

```{r keywords_overall, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
kp.df <- filter(pi.df, attributes == "KP")

@@

For this analysis (to-do), select a few keywords that may tell an interesting story of how they change over time -- something that might be revealing of 'trends.'

"social justice"
"evidence-based practice" | "evidence based practice" | "EBP"
"bullying"
"child welfare"

```{r keywords_over_time, warning=FALSE, message=FALSE, echo=FALSE}

@@

```{r location, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}

@@

print(location)
write.csv(location, "location.csv")

#img2 <- readPNG("/Users/beperron/Git/SocialWorkResearch/Chloro.png")
#grid.raster(img2)
```

@@

```{r location_over_time, warning=FALSE, message=FALSE, echo=FALSE}

@@

This section is currently not functional
```{r countries_not_represented, echo=FALSE}
#Tableau.countries <- read.csv("/Users/beperron/Git/SocialWorkResearch/SupportingDocs/Tableau\ Countries.csv")
```

@@

```{r methodology, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}

@@

```{r echo=FALSE}
year <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)
